---
title: "Activity Monitoring Data"
author: "Braydon"
date: "10/30/2019"
output: 
  html_document: 
    keep_md: yes
keep_md: true
---

Clear workspace
```{r}
rm(list=ls())
```
Load necessary libraries
```{r}
library(dplyr)
library(ggplot2)
```
Load the activity data file and explore
```{r}
activity_data <- read.csv("/Users/braydondavis/Downloads/activity.csv",header=TRUE,na.strings = "NA")
str(activity_data)
```
Change date from factor to a date
```{r}
activity_data_cleaned <- activity_data
activity_data_cleaned$date <- as.Date(activity_data_cleaned$date)
str(activity_data_cleaned)
```
Remove NAs from the data
```{r}
activity_data_cleaned <- na.omit(activity_data_cleaned)
```
Create new variables
```{r}
activity_data_cleaned$year <- as.numeric(format(activity_data_cleaned$date, "%Y"))
activity_data_cleaned$month <- as.numeric(format(activity_data_cleaned$date, "%m"))
activity_data_cleaned$day <- as.numeric(format(activity_data_cleaned$date, "%d"))
summary(activity_data_cleaned)
```
# Question 1 - What is mean total number of steps taken per day?
Create a histogram of the total number of steps taken per day
```{r}
steps_per_day <- activity_data_cleaned %>% select(steps, date) %>% group_by(date) %>% summarize(sum_steps = sum(steps))
ggplot(data = steps_per_day,aes(x=sum_steps))+
              geom_histogram(aes(y=..count..,fill=..count..),binwidth=1000,colour="black")+
              labs(x="Step Count",y="Frequency",title = "Frequency of Daily Step Count")
```
Find the mean total of number of steps taken
```{r}
mean(steps_per_day$sum_steps)
```
# Question 2 - What is the average daily activity pattern?
Create interval dataset with mean steps
```{r}
interval <- activity_data_cleaned %>% select(interval,steps) %>% group_by(interval) %>% summarize(mean_steps = mean(steps))
```
Create time series plot of average steps taken by interval
```{r}
ggplot(data = interval, aes(interval,mean_steps))+
        geom_line(color = "dark red")+
        ggtitle("Average Steps Taken by Interval")+
        labs(x= "Interval",y="Average Steps")
```
Find the interval that has the maximum number of steps
```{r}
interval[which.max(interval$mean_steps),]
```
# Question 3 - Impute missing values
Find the number of missing values in the dataset
```{r}
sum(is.na(activity_data$steps))
```
Impute NAs with mean 
```{r}
activity_data_filled <- activity_data
activity_data_filled <- activity_data_filled %>% group_by(interval) %>%
mutate(steps=ifelse(is.na(steps),mean(steps,na.rm=TRUE),steps))
```
Make a histogram of the total number of steps taken each day
```{r}
steps_per_day2 <- activity_data_filled %>% select (date,steps) %>% group_by(date) %>% summarize(steps = sum(steps))
ggplot(data=steps_per_day2,aes(x=steps))+
        geom_histogram(aes(y=..count..,fill=..count..),binwidth=1000,colour="black")
```
Calculate the mean of the full dataset without NAs
```{r}
mean(steps_per_day2$steps)
```
Calculate the median of the full dataset without NAs
```{r}
median(steps_per_day2$steps)
```
# Question 4 - Are there differences in activity patterns between weekdays and weekends?
Create Week_Type column
```{r}
activity_data_filled$date <- as.Date(activity_data_filled$date)
activity_data_filled <- activity_data_filled %>% mutate(week_type = ifelse(weekdays(date) == "Saturday" | weekdays(date) == "Sunday","Weekend","Weekday"))
```
Average steps taken in 5 minute intervals
```{r}
interval_full_avg <- activity_data_filled %>% group_by(interval,week_type) %>% summarize(steps = mean(steps))
```
Create a panel plot with geom_line
```{r}
ggplot(data = interval_full_avg, aes(x=interval,y=steps, color = week_type))+
        geom_line()+
        facet_wrap(~week_type,ncol=1,nrow=2)
```




